---
title: "Homework 2"
author: "Jacob Harrell"
date: "2/14/2021"
ouput: html_document
---

1. Calculate the residuals. Provide summary of residuals. Plot a histogram of the residuals. Do the residuals appear to be Gaussian?

```{r}
#setwd(choose.dir("New_Folder"))
beach <- read.csv(file='beach.csv')

plot(beach$Year, beach$OpenBeach, xlab='year', ylab='ha')

fit <- lm(formula= OpenBeach ~ Year * BeachID, data = beach)
summary(fit)

fit.res <- resid(fit)
summary(fit.res)

plot(beach$Year, fit.res, ylab = "residuals", xlab = "year", main = "Residuals for linear model")

summary(fit.res)

hist(fit.res, xlab = "Residual Value", main = "Residuals", col = "green", border = "black")

```


The residuals have certain qualities that may identify them as a Gaussian model. There is a distinct bell-shpaed curve when plotted on a histogram. This curve is much more defined to the left of the 0-point residual and weakly defined to the right of it. 

2. Calculate test statistics for regression coefficients. Verify by comparing test statistics reported from the model output.

```{r}
#setwd(choose.dir("New_Folder"))
beach<-read.csv(file='beach.csv')

fit <- lm(formula= OpenBeach ~ Year * BeachID, data = beach)

ts1<-coef(fit)[2]/summary(fit)[['coefficients']]['Year', 'Std. Error']
ts1

ts2<-coef(fit)[3]/summary(fit)[['coefficients']]['BeachIDB','Std. Error']
ts2

ts3<-coef(fit)[4]/summary(fit)[['coefficients']]['BeachIDC','Std. Error']
ts3

ts4<-coef(fit)[5]/summary(fit)[['coefficients']]['Year:BeachIDB','Std. Error']
ts4

ts5<-coef(fit)[6]/summary(fit)[['coefficients']]['Year:BeachIDC','Std. Error']
ts5

summary(fit)

```

3. Calculate p-values for regression coefficients. Verify values with model output. Waht are the associated null hypotheses? Do we reject or fail to reject these hypotheses?

```{r}
#setwd(choose.dir("New_Folder"))
beach<-read.csv(file='beach.csv')

fit <- lm(formula= OpenBeach ~ Year * BeachID, data = beach)

#p-value regression coefficient 1
pt(q=ts1, df=nrow(beach)-length(coef(fit)))*2
summary(fit)[['coefficients']]['Year','Pr(>|t|)']

#p-value regression coefficient 2
pt(q=ts2, df=nrow(beach)-length(coef(fit)))*2
summary(fit)[['coefficients']]['BeachIDB','Pr(>|t|)']

#p-value regression coefficient 3
pt(q=ts3, df=nrow(beach)-length(coef(fit)))*2
summary(fit)[['coefficients']]['BeachIDC','Pr(>|t|)']

#p-value regression coefficient 4
pt(q=ts4, df=nrow(beach)-length(coef(fit)))*2
summary(fit)[['coefficients']]['Year:BeachIDB','Pr(>|t|)']

#p-value regression coefficient 5
pt(q=ts5, df=nrow(beach)-length(coef(fit)))*2
summary(fit)[['coefficients']]['Year:BeachIDC','Pr(>|t|)']

```

In each instance, the null hypotheses for the regression coefficients is that the coefficient associated with the respective variable at 'year', 'beachIDB', 'beachIDC', 'year:BeachIDB', and 'year:beachIDC'equals zero. Because each respective p-value is smaller than any reasonable a, in each instance the null hypotheses would be rejected.

4. Select a single regression coefficient. Devise a null hypothesis different from the default. Report test statistics, p-value, and whether the null hypothesis is rejected or fails to reject. 

In this instance, I have selected the regression coefficient 'BeachIDB' for further analysis.The null hypothesis in this case is that BeachIDB=100. 

```{r}
#setwd(choose.dir("New_Folder"))
beach<-read.csv(file='beach.csv')

fit <- lm(formula= OpenBeach ~ Year * BeachID, data = beach)

ts_new<-(coef(fit)[3]-100)/summary(fit)[['coefficients']]['BeachIDB','Std. Error']
ts_new
pt(q=ts_new, df=nrow(beach)-length(coef(fit)))*2
summary(fit)[['coefficients']]['BeachIDB','Pr(>|t|)']

```
We would reject the null hypothesis because the associated p-value is smaller than any reasonable a. 

5. Interpret output of the fitted model. How does the beach area change through time at each of the three beaches?

```{r}

#setwd(choose.dir("New_Folder"))
beach<-read.csv(file='beach.csv')

fit <- lm(formula= OpenBeach ~ Year * BeachID, data = beach)

change<-coef(fit)
#beach A
change[2]
#beach B
change[2]+change[5]
#beach C
change[2]+change[6]

```

Based on the fitted model, there is evidence that each beach experiences a statistical decline in open beach area.
Beach A has an annual decline of 18.36491 hectares.
Beach B has an annual decline of 1.322145 hectares.
Beach C has an annual decline of 3.993592 hectares. 


