---
title: "Homework 1 WMAN"
author: "Jacob Harrell"
date: "2/10/2021"
output:
  html_document: default
---

1. Import Data Set and Examine First Few Rows
```{r}
read.csv('beach.csv')

```

```{r}
beach <- read.csv('beach.csv')
head(beach)
```

2. Fit a model to include interactions between Year and BeachID. Report estimates of all Coefficients. 

```{r}
plot(beach$Year, beach$OpenBeach, xlab='year', ylab='ha')

fit <- lm(formula= OpenBeach ~ Year * BeachID, data = beach)

x_new <- seq(from = 1992, to = 2014, length.out = 100)
y_new <- 36945.7-18.36491*x_new

lines(x = x_new, y = y_new)

```

Linear Model : y = -18.36491x + 36945.7
Coefficients:

Intercept = 36945.73
Year = -18.36491
BeachIDB = -34211.98
BeachIDC = -28776.11
Year:BeachIDB = 17.04277
Year:BeachIDC = 14.37132

3. Predict the beach area at all three beaches in 1992. 

```{r}
x <- c(1992)
y <- 36945.7 - 18.36491 * x
```

Using the linear model, we can predict that the open beach area across all three beaches in 1992 was 362.8 ha.

4. Plot the predicted area of beach across all years.

BeachA
```{r}
beachA <- read.csv(file = 'beachA.csv')

plot(beachA$Year, beachA$OpenBeach, xlab='year', ylab='ha')

fitA <- lm(formula = OpenBeach ~ Year, data = beachA)
View(fitA)

x_A <- seq(from = 1992, to = 2014, length.out = 100)
y_A <- 36945.7-18.36491*x_A

lines(x = x_A, y = y_A)

```

BeachB

```{r}
beachB <- read.csv(file='beachB.csv')

plot(beachB$Year, beachB$OpenBeach, xlab='year', ylab='ha')

fitB <- lm(formula = OpenBeach ~ Year, data = beachB)
View(fitB)

x_B <- seq(from = 1992, to = 2014, length.out = 100)
y_B <- 2733.746-1.322145*x_B

lines(x = x_B, y = y_B)

```

BeachC

```{r}

beachC <- read.csv(file='beachC.csv')

plot(beachC$Year, beachC$OpenBeach, xlab='year', ylab='ha')

fitC <- lm(formula = OpenBeach ~ Year, data = beachC)
View(fitC)

x_C <- seq(from = 1992, to = 2014, length.out = 100)
y_C <- 8169.61-3.993592*x_C

lines(x = x_C, y = y_C)

```

5. Interpret the rate of beach area change at each of the three beaches.

The rate of change across each of the three beaches is -18.36491. That is, the amount of open beach area decreases by a rate of 18.36491 hectares per year. Taking a look at the rate of change across each beach individually, we find that beach A decreased at a rate of 18.36491 hectares per year, beach B decreased at a rate of 1.322145 hectares per year, and beach C decreased at a rate of 3.993592 hectares per year.


