---
title: "Homework 5"
author: "Jacob Harrell"
date: "3/30/2021"
output: html_document
---
```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

<b>Question 1</b>

First load the data and then place into an unmarkedFrameOccu object.

```{r}
setwd("C:/Users/Jacob/Desktop/HW5")
bob<-read.csv(file = 'Bobcat.csv')

library(unmarked)

bob_mat <- as.matrix(bob)
occu_data <- unmarkedFrameOccu(y = bob_mat)
```

<b>Question 2</b>

Fit an occupancy model with the assumption that conditional detection probability is a function of the number of people photographed at a site and that the occupancy probability is a function of disturbances within 5 km of a site. 

```{r}
p_covs <- read.csv("p covariates.csv")

det_covs <- list(people = data.frame(p_covs))

site_covs <- read.csv('psi covariates.csv')

occu_data2 <- unmarkedFrameOccu(y = bob_mat, siteCovs = site_covs, obsCovs = det_covs)
fit <- occu(formula = ~ people ~ Dist_5km, data = occu_data2)
summary(fit)
```

<b>Question 3</b>

Interpret the effect of the number of people photographed at each site on the conditional detection probability of bobcat. Verify through r.


```{r}
betas <- coef(fit)
betas

p_1 <- plogis(betas[3]+betas[4]*1)
p_1
p_2 <- plogis(betas[3]+betas[4]*2)
p_2

log((p_2/(1-p_2))/(p_1/(1-p_1)))

```
For every one unit change within the variable 'people', the log odds of success of detecting a bobcat decreases by .0609. 

<b>Question 4</b>

Re-create the Wald test for the effect of disturbance within 5 km. Define the null hypothesis and note whether we would reject it or fail to reject it. 

```{r}

disturb <- -23.64 / 4.773 #coefficient for dist_5km divided by its standard error
disturb

2 * pnorm(-1 * abs(disturb))
```
The null hypothesis, plainly stated, is that the log odds of success of detecting a bobcat at a site would not differ depending on the proportion of the site area within 5km of a recently disturbed area. Based on the results of the Wald Test, we would reject the null hypothesis. 

<b>Question 5</b>

Predict and plot the effect of disturbance within 5 km on bobcat occupancy probability over the range of observed disturbance within 5 km. 

```{r}
new_psi <- data.frame(Dist_5km = seq(from = min(site_covs$Dist_5km),
                                to = max(site_covs$Dist_5km),
                                length.out = 100)
)

prd <- predict(object = fit, newdata = new_psi, type = 'state')
head(prd)
```

```{r}

plot(y= prd$Predicted, x = new_psi$Dist_5km, 
     ylab = 'Bobcat Occupancy Probability', 
     xlab = 'range of observed disturbance over 5km area',
     cex.axis = 1.5,
     cex.lab = 1.5,
     type = 'l')
lines(y = prd$lower, x = new_psi$Dist_5km, lty = 2)
lines(y = prd$upper, x = new_psi$Dist_5km, lty = 2)
```

